---
title: "Snore"
output: html_document
date: "2023-03-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Snore Project
##Introduction

The aim of the project is to analyze the factors that are effecting sleep efficiency. The data set
is one of the Kaggle dataset and called "Sleep Efficiency Dataset". The dataset provides data on a group of test volunteers and their sleeping habits. Each subject has a "Subject ID" and "Sleep efficiency" feature is a measure of the proportion of time spent in bed that is actually spent asleep.
The data also contains information on each subject's caffeine and alcohol use in the 24 hours before bedtime, smoking status, and exercise frequency.

###Research Question
How do people's lifestyle and habits affect the length and quality of their sleep?
Does gender also have an effect on sleep quality?
###Hypotheses
People who consume more alcohol, caffeine and tobacco are likely to sleep less efficiently.
Conversely, people who exercise more are likely to sleep more efficiently.


```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
