---
title: "Snore"
date: "2023-03-11"
output: 
    html_document:
      toc: true
      toc_float: true
      theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("tidyverse")
#nstall.packages("dplyr")
#install.packages("BBmisc")
library("rio")
library("dplyr")
library("tidyverse")
library("ggplot2")
library("BBmisc")
library("stargazer")
library("here")
# ------- 1. Importing and Cleaning
# Import data
sleep_data <-import(here("Sleep_Efficiency.csv"))
```

# Snore Project
## Introduction

The aim of the project is to analyze the factors that are effecting sleep efficiency. The data set
is one of the Kaggle dataset and called "Sleep Efficiency Dataset". The dataset provides data on a group of test volunteers and their sleeping habits. Each subject has a "Subject ID" and "Sleep efficiency" feature is a measure of the proportion of time spent in bed that is actually spent asleep.
The data also contains information on each subject's caffeine and alcohol use in the 24 hours before bedtime, smoking status, and exercise frequency.

### Research Question
- How do people's lifestyle and habits affect the length and quality of their sleep?
- Does gender also have an effect on sleep quality?

### Hypotheses
- People who consume more alcohol, caffeine and tobacco are likely to sleep less efficiently.
- Conversely, people who exercise more are likely to sleep more efficiently.

### 2. Data Visualisation 1: Impact of caffeine consumption on sleep efficiency

```{r, warning=FALSE}
library("dplyr")
# Histogram of age distribution
histogram_caffeine <- sleep_data %>% 
  na.omit(`Caffeine consumption`) %>% 
  ggplot(aes(`Caffeine consumption`))+
  geom_histogram(stat="count")
histogram_caffeine
# Divide caffeine consumption into groups:
# None=0
# Low+ 25
# Medium=50
# High >=75
sleep_data_caffeine_groups <- sleep_data %>% 
  na.omit(`Caffeine consumption`) %>% 
  mutate(caffeine_group=case_when(`Caffeine consumption`==0 ~ "None",
                                  `Caffeine consumption` ==25 ~ "Low (25mg)",
                                  `Caffeine consumption` ==50 ~ "Medium (50mg)",
                                  TRUE ~ "High (75mg-200mg)"))
# Convert caffeine consumption groups column to factor variable 
sleep_data_factor <- sleep_data_caffeine_groups %>% 
  mutate(caffeine_group= factor(caffeine_group,
                                ordered=TRUE,
                                levels=c("None","Low (25mg)","Medium (50mg)","High (75mg-200mg)")))
# Box plots indicating how caffeine consumption affects sleep efficiency
caffeine_sleep_efficiency <- ggplot(sleep_data_factor, aes(x=caffeine_group, y=`Sleep efficiency`)) + 
  geom_boxplot()+
  xlab(" Amount of Caffeine Consumed")+
  ylab("Sleep Efficiency")+
  ggtitle("Effect of Caffeine Consumption on Sleep Efficiency")+
  theme_minimal()
caffeine_sleep_efficiency
```
Box plots indicating how caffeine consumption affects sleep efficiency.
Counter intuitively appears that drinking more coffee improves sleep efficiency i.e amount of time in bed spent asleep, but is this difference statistically significant? The median
sleep efficiency in the high group is larger than the others, but all groups overlap.

Why could this be? Do people who drink more coffee, exercise more which could lead to more efficient sleep? 

Do they drink more/less alcohol? Is gender an important factor? Could caffeine consumption affect the type or duration of sleep?


### 2. Data Visualisation 2: Impact of gender and caffeine consumption on sleep efficiency


```{r}
# Plot using the same code as above but split by gender
caffeine_gender_sleep_efficiency <- ggplot(sleep_data_factor, aes(x=caffeine_group, y=`Sleep efficiency`, fill=Gender)) + 
  geom_boxplot()+
  xlab(" Amount of Caffeine Consumed")+
  ylab("Sleep Efficiency")+
  ggtitle("Effect of Caffeine Consumption and Gender on Sleep Efficiency")+
  theme_minimal()
caffeine_gender_sleep_efficiency
```

Seems to be a difference in the median sleep efficiency between the genders for each level of caffeine consumption, but is it significant? For women consuming no or high levels of caffeine seems to increase sleep efficiency, while low to medium amounts decrease it.(Quadratic shape?) 

Mens' sleep efficiency seems to increase with the amount of caffeine consumed. Seems unlikely caused by caffeine affecting men and women in different ways, instead likely to have different lifestyles.

### 2. Data Visualisation 3. Comparison of lifestyle choices between genders
```{r,warning=FALSE}
# Histogram of exercise frequency by gender
histogram_exercise <- sleep_data %>% 
  drop_na(`Exercise frequency`) %>% 
  ggplot(aes(`Exercise frequency`, fill=Gender))+
  geom_histogram(stat="count", position="dodge")+
  ggtitle("Difference in exercise frequency between men and women")
histogram_exercise
# Histogram of alcohol consumption by gender
histogram_alcohol <- sleep_data %>% 
  drop_na(`Alcohol consumption`) %>% 
  ggplot(aes(`Alcohol consumption`, fill=Gender))+
  geom_histogram(stat="count", position="dodge")+
  ggtitle("Difference in alcohol consumption between men and women")
histogram_alcohol
```
Men seem to exercise considerably more than women.
Not big difference in alcohol consumption between genders, most participants in sleep study don't rink at all

## Data Modelling

After seeing which factors lead to better or worse sleeping efficiency, the next step would be to fit a data model to predict new data points. We will include a rudimentary linear regression model.

```{r}
#Data modeling
library(BBmisc)
library(stargazer)
sleep <- read.csv(here("Sleep_Efficiency.csv"))
sl1 <- sleep %>% 
  select(c(ID, Sleep.efficiency, Caffeine.consumption, Alcohol.consumption, Smoking.status, 
           Exercise.frequency, Awakenings))
slsc <- normalize(x = sl1[,2:7])
summary(slsc)
```

```{r}
# Get a first look at data
head(slsc, n = 5)
str(slsc)
```

Before modelling, we need to clean the data. A first step for this is finding NAs and replacing them with the mean of the variable in question.
```{r}
# Check for missing values

cat("Number of missing values:\n")
cat("by column:\n")
nan_val_count <- colSums(is.na(slsc))
print(nan_val_count)

som <- sum(nan_val_count)
cat("Total:", som, " ~ ", round(som / (nrow(slsc) * ncol(slsc)) * 100), "% of the dataset\n")

# replace null values with mean

library(dplyr)

slsc$Awakenings[is.na(slsc$Awakenings)] <- mean(slsc$Awakenings, na.rm = TRUE)
slsc$Caffeine.consumption[is.na(slsc$Caffeine.consumption)] <- mean(slsc$Caffeine.consumption, na.rm = TRUE)
slsc$Alcohol.consumption[is.na(slsc$Alcohol.consumption)] <- mean(slsc$Alcohol.consumption, na.rm = TRUE)
slsc$Exercise.frequency[is.na(slsc$Exercise.frequency)] <- mean(slsc$Exercise.frequency, na.rm = TRUE)
slsc$Sleep.efficiency[is.na(slsc$Sleep.efficiency)] <- mean(slsc$Sleep.efficiency, na.rm = TRUE)
```
### Linear regression model

```{r}
model1<-lm(slsc$Sleep.efficiency~slsc$Caffeine.consumption+slsc$Alcohol.consumption+slsc$Smoking.status+slsc$Exercise.frequency)
summary(model1)
summary(model1)$coefficient
confint(model1)
```
```{r, results='asis'}
stargazer(model1, type='html')
```




## Results & Conclusion

In this exercise, we analyzed the efect of certain habits on people's sleeping quality. We used multiple regression analysis in our model and tried to understand the effect of alcohol, tobacco and caffeine consumption and exercise frequency on sleep efficiency. 

The results of the regression table show that alcohol and tobacco consumption and exercise frequency are factors have statistically significant impact on sleep efficiency(p>0.05). However, caffeine consumption and gender had no statistically significant effect on sleep efficiency.  

People who consume alcohol are 0.03 times less likely to have high sleep efficiency. As well as smokers have 0.07 times less likely to have high sleep efficiency.

On the contrary, as the frequency of exercise increases, sleep quality also increases. People who exercise frequently 0.02 times more likely to have high sleep efficiency. 

In conclusion, our analyses showed that certain habits effect sleep efficiency. Although we were expecting that caffeine consumption would effect the sleep efficiency, we couldn't find any significant importance. Also, gender difference has no impact on sleep efficiency.

Alcohol and tobacco consumption have negative effect on sleeping quality. Yet, exercise frequency increases the sleep quality.